---
title: "Version-4.0-New-Features"
author: "Zach Deane-Mayer"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Version-4.0-New-Features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

caretEnsemble 4.0.0 introduces many new features! Let's quickly go over them.

# Multiclass support
caretEnsemble now fully supports multiclass problems:
```{r}
set.seed(42)
model_list <- caretEnsemble::caretList(
  x = iris[, 1:4],
  y = iris[, 5],
  methodList = c("rpart", "rf")
)
```

# Greedy Optimizer in caretEnsemble
The new version uses a greedy optimizer by default, ensuring the ensemble is never worse than the worst single model:
```{r}
ens <- caretEnsemble::caretEnsemble(model_list)
summary(ens)
```

# Enhanced S3 Methods
caretStack (and by extension, caretEnsemble) now supports various S3 methods:
```{r}
print(ens)
summary(ens)
plot(ens)
ggplot2::autoplot(ens)
```
# Improved Default trainControl
A new default trainControl constructor makes it easier to build appropriate controls for caretLists. These controls include explicit indexes based on the target, return stacked predictions, and use probability estimates for classification models.
```{r}
class_control <- caretEnsemble::defaultControl(iris$Species)
print(ls(class_control))
```

```{r}
reg_control <- caretEnsemble::defaultControl(iris$Sepal.Length)
print(ls(reg_control))
```

# Mixed Resampling Strategies
Models with different resampling strategies can now be ensembled:
```{r}
set.seed(42)
y <- iris[, 1]
x <- iris[, 2:3]
flex_list <- caretEnsemble::caretList(
  x = x,
  y = y,
  methodList = c("rpart", "rf"),
  trControl = caretEnsemble::defaultControl(y, number = 3)
)

flex_list$glm_boot <- caret::train(
  x = x,
  y = y,
  method = "glm",
  trControl = caretEnsemble::defaultControl(y, method = "boot", number=50)
)

flex_ens <- caretEnsemble::caretEnsemble(flex_list)
print(flex_ens)
```

# Mixed Model Types
caretEnsemble now allows ensembling of mixed lists of classification and regression models:
```{r}
set.seed(42)
X <- iris[,1:4]
target_class <- iris[, 5]
target_reg <- as.integer(iris[, 5] == 'virginica')
model_class <- caret::train(iris[, 1:4], target_class, method = "rf", trControl = caretEnsemble::defaultControl(target_class))
model_reg <- caret::train(iris[, 1:4], target_reg, method = "rf", trControl = caretEnsemble::defaultControl(target_reg))
mixed_list <- caretEnsemble::as.caretList(list(class=model_class, reg=model_reg))
mixed_ens <- caretEnsemble::caretEnsemble(mixed_list)
print(mixed_ens)
```

# Transfer Learning
caretStack now supports transfer learning for ensembling models trained on different datasets:
```{r}
set.seed(42)
train_idx <- sample(1:nrow(iris), 100)
train_data <- iris[train_idx, ]
new_data <- iris[-train_idx, ]

model_list <- caretEnsemble::caretList(
  x = train_data[, 1:4],
  y = train_data[, 5],
  methodList = c("rpart", "rf")
)

transfer_ens <- caretEnsemble::caretEnsemble(
  model_list,
  new_X = new_data[, 1:4],
  new_y = new_data[, 5]
)

print(transfer_ens)
```

We can also predict on new data:
```{r}
preds <- predict(transfer_ens, newdata = head(new_data))
print(preds)
```

# Permutation Importance
Permutation importance is now the default method for variable importance in caretLists and caretStacks:
```{r}
importance <- caret::varImp(transfer_ens)
print(importance)
```

This completes our demonstration of the key new features in caretEnsemble 4.0. These enhancements provide greater flexibility, improved performance, and easier usage for ensemble modeling in R.